---
title: "ESM 244 Lab 1"
author: "Jaxon Stuhr"
date: "1/4/2022"
output: html_document
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Read in data

```{r}
sf_trees = read_csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Part 1: Warngling and ggplot

**Example 1:** Find counts of observation by `legal_status` and wrangle.

```{r}
### method 1: group_by() and summarize()
sf_trees %>% 
  group_by(legal_status) %>% 
  summarize(tree_count = n())

### method 2: count
top_5_status = sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% # moves variable to front
  slice_max(tree_count, n = 5) %>% 
  arrange(-tree_count)
```

Make a graph of top 5 from above:

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) + 
  geom_col(fill = 'darkgreen') +
  labs(x = "Legal Status", y = "Tree Count") +
  coord_flip() + # switch axes
  theme_minimal()
```

**Example 2:** Only going to keep observations where legal status is "Permitted Site" and caretake is "MTA", and store as `permitted_data_df`

```{r}
permitted_data_df = sf_trees %>% 
  filter(legal_status %in% c("Permitted Site", "Private"), caretaker == "MTA")
```

**Example 3:** Only keep blackwood acacia trees, and then ol keep columns `legal_status`, `date`, `latitude`, `longitude` and store as `blackwood_acacia_df`

```{r}
blackwood_acacia_df = sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, lat = latitude, long = longitude)

# make a graph of locations
ggplot(data = blackwood_acacia_df, aes(x = long, y = lat)) +
  geom_point(color = 'darkgreen')
```

**Example 4:** Separate function

```{r}
sf_trees_sep = sf_trees %>% 
  separate(species, into = c('spp_scientific', 'spp_common'), sep = ' :: ')
```


**Example 5:** Use reunite function

```{r}
ex_5 = sf_trees %>% 
  unite('id_status', tree_id, legal_status)
```

